### 2004-07-12 Documentation/part2.tex
Some changes in part II: simbinXXX -> samplingXXX Type 2 smearing algorithm decsription replaced by section 'Low pass filter' Checked that the commands in the documentation are the same than in the code.

\subsection{Low-pass filter based automatic energy smearing}

THIS SUBSECTION WILL BE RE-WRITTEN BY PATRICK (WW)

The second available algorithm is more complicated, a little slower but
much more accurate and much less susceptible to aliasing effects
\index{Aliasing}. The starting
point again is equation~\ref{eq:simple_int}. The idea now is to
 have a good approximation to the integral but still keeping the fixed sampling
point in $E$. To this end we rewrite the integrand of 
equation~\ref{eq:simple_int} using the sampling theorem~\cite{NRC,Rybicki}
\begin{equation}
\label{eq:sampling}
f(E)=\sum_{k=-\infty}^{\infty}\,f(E_k) 
\times \mathrm{sinc}\left(\frac{\pi}{h}(E-E_k)\right)+e(E)
\end{equation}
with $\mathrm{sinc}(x):=\sin(x)/x$ and $E_k=E^0+h\cdot k$. 
This the so called sampling 
representation of the integrand $f(E)$ and $e(E)$ is an error term.
The error term vanishes if the Fourier transform $\hat f(\omega)$ of $f(E)$
is zero outside the interval $[-\pi/h,\pi/h]$. Thus one suspects that
the error term is small once the power beyond $\pi/h$ is low. Furthermore
the sum in equation~\ref{eq:sampling} can be truncated to a few terms if
$f(E)$ has a compact support, \ie\ it is different from zero only 
on a finite interval. Thus the sampling representation is most useful
for functions which have both a compact support in real \emph{and} Fourier
space. This of course is impossible, however the function which comes
most closely to this, is well known -- it is a Gau\ss ian. Our $f(E)$ is not
exactly a Gau\ss ian  but the convolution of a top hat and a Gau\ss ian, thus
its real and Fourier space components still behave rather Gau\ss ian 
especially for large values of $E$ and $\omega$ respectively. The advantage of
using the sampling representation is that the integration with respect to $E$
is easily performed analytically
  \begin{eqnarray}
\label{eq:int_sampling}
\int dE \quad f(E)&\simeq& \sum_{k=k_{l}}^{k_{u}}\,f(E_k) 
\times \mathrm{Si}\left(\frac{\pi}{h}(E-E_k)\right)\,,\\
\mathrm{Si}(x)&:=&\int_{0}^x dx'\quad \sin(x')/x'\,,\nonumber 
\end{eqnarray}
where $k_l$ and $k_u$ are suitable chosen bounds for the summation.

In order to have a good compromise between speed an accuracy the choice
of $h$, $k_l$ and $k_u$ is crucial. The choice of $h$ is governed by the
need to sample $f(E)$ at a sufficient rate in order not to produce aliasing.
For this purpose it useful to consider some properties of the power spectrum
of $f(E)$. Provided the cross section and flux is smooth enough  the only
high power components in $\hat f(\omega)$ are due to a fast oscillating
$P(E)$. We can estimate the largest frequency $\omega_m$ quite easily, it is
given by the largest $\Delta m^2_m$ and the lowest energy $E_l$
\begin{equation}
\omega_m \simeq \frac{\Delta m^2_m \cdot L}{E_l^2}
\end{equation}
On the other hand the finite energy resolution of any real detector anyhow
will act as a low-pass filter eliminating any power beyond $\sim 1/\sigma(E)$.
For now we assume that there are no frequencies which are not resolved by
the detector thus the sampling frequency is determined by the energy 
resolution of the detector. In that case the correct sampling 
width $h$ is approximately
\begin{equation}
h\simeq 1/\sigma(E_l)\,.
\end{equation}

The range for the summation $k_l^i$ and $k_u^i$ is given by the range of 
integration $a_i$ and $b_i$ in equation~\ref{eq:simple_int}, the integration 
limits there are $0$ to $\infty$. For practical purposes however the range
can be reduced to $a_i$ and $b_i$. They are  determined by requiring
that the area under the bin kernel $K_i^c$ has reached a specific 
value close to one,
the so called confidence level $c$ (new name would be better). $1-c$ is a 
small number and specifies the fraction of  events in bin which lie outside
the summation range, usually $1-c=10^{-3}$ is sufficient. Furthermore the
value of the bin kernel at the lower and upper boundary should
be the same
\begin{eqnarray}
\int_{a_i}^{b_i} dE\quad K_i^c(E) &=& c\,,\\
K_i^c(a_i)&=&K_i^c(b_i)\,.
\end{eqnarray} 
Knowing $a_i$ and $b_i$ it is easy to obtain $k_l$ and $k_u$ by requiring
that the summation extends one bin beyond  each $a_i$ and $b_i$. Furthermore
it is possible to specify an offset which extends the index range on each side
by its value.

Since the $a_i$'s and $b_i$'s are not known in advance it could be possible
that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
\eg\ no cross section or flux is defined at those energies. To remedy this 
problem the computational domain can be specified. No calculation in \GLOBES\
will leave this domain.

In order to ensure that fast oscillating probabilities do not lead to aliasing
it is desirable to impose a low-pass filter already during the calculation
of the probabilities itself. This is possible and implemented has a highly
experimental feature called ``filter''\index{Filter}. 
The calculation of oscillation
probabilities basically is a computation of phase differences. Restricting
the maximum admissible size of those phase difference effectively filters
the high frequency component of the oscillation probability. This idea is
implemented according to
\begin{eqnarray}
\label{eq:filter_a}
P_{\alpha\beta}(E)&=&\sum_{ij}
U_{\alpha j} U^*_{\beta j} U^*_{\alpha i} U_{\beta i} 
e^{-i\Phi_{ij}}\times 
e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
\end{eqnarray}
where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
approximately like an energy resolution function with constant width 
$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
\begin{equation}
\label{eq:filter_b}
\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
\end{equation}
The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
is not obvious and connected to the properties of $P_{\alpha\beta}$, 
see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
for vacuum and constant densities and controlled
by the filer state variable and $\sigma_f^0$ is set by the filter value 
variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
frequencies one can split the energy resolution function $\sigma(E)^c$ in
two parts by
\begin{equation}
\sigma_c^2(E)=\underbrace{\sigma_c^2(E)-\sigma_c^2(E_\mathrm{min})}_
{\tilde\sigma^2_c(E)}+\underbrace{\sigma_c^2(E_\mathrm{min})}_{\sigma_e^2}\,,
\end{equation}
where $\tilde\sigma_c(E)$ is used instead of $\sigma_c(E)$ in computing the
smearing data and $\sigma_e$ is used to determine $\sigma_f^0$.

>>>>

\subsection{Low-pass filter}

In order to ensure that fast oscillating probabilities do not lead to 
aliasing\index{Aliasing}
it is desirable to impose a low-pass filter already during the calculation
of the probabilities itself. This is possible and implemented has a highly
experimental feature called ``filter''\index{Filter}. 
The calculation of oscillation
probabilities basically is a computation of phase differences. Restricting
the maximum admissible size of those phase difference effectively filters
the high frequency component of the oscillation probability. This idea is
implemented according to
\begin{eqnarray}
\label{eq:filter_a}
P_{\alpha\beta}(E)&=&\sum_{ij}
U_{\alpha j} U^*_{\beta j} U^*_{\alpha i} U_{\beta i} 
e^{-i\Phi_{ij}}\times 
e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
\end{eqnarray}
where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
approximately like an energy resolution function with constant width 
$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
\begin{equation}
\label{eq:filter_b}
\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
\end{equation}
The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
is not obvious and connected to the properties of $P_{\alpha\beta}$, 
see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
for vacuum and constant densities and is controlled
by the filer state variable and $\sigma_e$ is set by the filter value 
variable:
\begin{quote}
{\tt
\$filter\_state = 1\\
\$filter\_value = 2.0\\
}
\end{quote}
would switch the filter feature on and set the width to $2.0\,\mathrm{GeV}$.
The setting of {\tt \$filter\_state} is ignored whenever a density profile
with more than one layer is used.

With a type 1 ({\tt @type = 1}) energy resolution 
function, $\sigma_e$ adds on to the energy resolution function 
of the detector $\sigma_c(E)$ like
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_e^2 + \sigma_c(E)^2\,.
\end{equation} 
Sometimes this behaviour is undesirable and therefore one can try to 
'subtract' the filtering from the energy resolution function by splitting 
the energy resolution function $\sigma(E)_{\mathrm{eff}}$ in
two parts by
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2=\underbrace{\sigma_c(E)^2-\sigma_e^2}_
{\tilde\sigma^2_c(E)}+\sigma_e^2\,,
\end{equation}
where the truncated energy resolution function $\tilde\sigma_c(E)$ 
is used instead of $\sigma_c(E)$ in computing the
smearing data. Thus one obtains that
\begin{equation}
\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_c(E)^2\,.
\end{equation} 
This scheme is enabled by choosing as type for the energy resolution
\begin{quote}
{\tt
@type = 2
}
\end{quote}
